Growing up with Nell: A Narrative Interface for Literacy

### 2012-03-08 paper.tex
Finish 1st draft of 'Narrative Interface' section under 'Key Ideas.

Children (like all humans) are hard-wired for
stories~\cite{boyd:stories}.  Nell uses a
\textit{Narrative Interface}~\cite{bizzocchi:narrative,don:narrative,laurel:computers} 
to guide the child user through the pedagogical capabilities of the system.

% excerpt of story model
% not always a story -- kid can draw on their own and return to the
% story; Nell will observe the free play and offer suggestions.
\subsection{Personal}
% customization, renaming characters
\subsection{Direct interaction}
% handwriting as UI?
% motor skills
% photo of XO-3?
\subsection{Growable (``no ceiling'')}
% it has a ``low floor and no ceiling'' to allow a child to master more
% and more of the system as she grows.

@@

\section{Technology}
\subsection{Rule-based story modules}
% rete algorithm, cite Riedl, etc.
% programmable conflict resolution
\subsection{Simple dialog with AIML}
% eliza interactions, show fragment
\subsection{Web technologies}
% Javascript/HTML5
% offline caching
% turtles (almost) all the way down (turtlescript again)

>>>>

Children (like all humans) are hard-wired for
stories~\cite{boyd:stories}.  Nell uses a
\textit{Narrative Interface}~\cite{bizzocchi:narrative,don:narrative,laurel:computers} 
to guide the child user through the pedagogical capabilities of the
system.  All system actions are shaped by the storybook metaphor, and
interactions with the system are conceptualized as interactions with
one of the system protagonists.

The overall system story is a multi-character serial adventure, taking
cues from \textit{The Diamond Age} and
from the UNIVERSE narrative-generation system \cite{lebowitz:universe}.
Each character represents a specific skill or subject area, and each
adventure represents about a year's curriculum.  This design also
improves modularity and allows updates and decentralized authorship.

As shown in figure~\ref{fig:nell}, Nell's characters are
always-available agents running on top of a system ``activity''
(application) which provides specialized functionality.  The agents
are not always foregrounded; the child is always free to engage in
constructionist play within the base activity.  The handwriting tutor is
fundamentally a drawing activity; the adventure involving the magical
musical lock is also a music-making activity.  The narrative system is
hooked into the activity to provide passive guidance
(congratulating the child when it notices they've drawn a letter),
active guidance (Apple-Guide--style~\cite{powers:appleguide}
contextual help), or system services (switch activities; jump into a
related story module/lesson plan).

% Story Module == Lesson Plan

% not always a story -- kid can draw on their own and return to the
% story; Nell will observe the free play and offer suggestions.
\subsection{Personal}
% customization, renaming characters

Kids love to personalize stuff \textbf{[CITATION NEEDED]}.

\subsection{Direct interaction}
% handwriting as UI?
% motor skills
% photo of XO-3?
\subsection{Growable (``no ceiling'')}
% it has a ``low floor and no ceiling'' to allow a child to master more
% and more of the system as she grows.

@@

\section{Technology}
\subsection{Rule-based story modules}
% excerpt of story model
% rete algorithm, cite Riedl, etc.
% programmable conflict resolution
\subsection{Simple dialog with AIML}
% eliza interactions, show fragment
\subsection{Web technologies}
% Javascript/HTML5
% offline caching
% turtles (almost) all the way down (turtlescript again)
